SimBio: a flexible and extensible Python package for systems modelling

### 2023-11-10 article.qmd
Write discussion

We have written a set of Python packages
that allow to define and simulate dynamical systems and reaction networks.
Thay have a modular architecture that makes them easy to extend in different points.

Poincaré and SimBio are focused on their integration with IDE

Modular and extensible set of python libraries
Leverage IDE integration and static typing for model creation

Introduce 3 layers: symbolite, poincare and simbio
easier to understand and develop
usable, extendable, and maintainable by a broader community
protocols
change in the future

>>>>

In this article,
we have described a set of Python packages we developed
that allow to define and simulate dynamical systems and reaction networks.
They are focused on a deep integration with IDEs
such that code analysis tools can flag errors before running the code,
and help with refactoring and code completion.
For this,
we chose to use standard modern Python syntax,
since the integration with IDEs will be maintained by the big players in the huge Python community.

In contrast to previous tools,
both model definition and compilation into an ODE function is implemented in Python.
This lowers the barrier to develop different simulation modalities,
such as perfomance optimizations by leveraging some structure in the model.
But being in Python does not mean less performance than C/C++ tools,
as the resulting ODE function can be *jitted* using numba.

These set of packages have a modular architecture with a clear separation of concerns.
As such, they can be reused, improved and extended by the broader Python community.
For example, people outside the systems biology community could
add a stochastic integrator to poincaré,
which could then be used from SimBio.
This architecture also makes them easier to understand,
reducing the difficulty to contribute fixes or new code,
assuring their maintainance and continued development into the future.
